---
title: "tiler"
output:
  rmarkdown::html_document:
    toc: false
    fig_caption: true
    self_contained: yes
fontsize: 11pt
documentclass: article
vignette: >
  %\VignetteIndexEntry{tiler}
  %\VignetteEngine{knitr::rmarkdown_notangle}
---

## Introduction

The _tiler R_ package can be used to render spatial data sets as [tiles](https://en.wikipedia.org/wiki/Tile_Map_Service). The advantage of rendering spatial data as tiles for use in web applications is that the tile format reduces the amount of data that users need to download to view the data. The disadvantage, however, is that the spatial data first needs to be converted to the tile format. Depending on the size of your data and the computational resources that you have available, tiling your data could take a very long time. I would recommend trying out the normal mapping functions available in the [_leaflet R_ package](https://rstudio.github.io/leaflet/), and make sure that you really need to use tiling before out trying out this package.

## Installation

The _tiler R_ package requires _python_ (version 2.7+) installed. It also requires that the _gdal python_ package is also installed. I have found setting up this software on Windows 7 to be quite a nightmare. Therefore, the _tiler R_ package is only available on Unix systems for the moment. The easiest way to setup a system on Ubuntu (version 16+) is using the following system commands.

```{bash, eval = FALSE}
sudo apt-add-repository ppa:ubuntugis/ubuntugis-unstable
sudo apt-get update
sudo apt-get install gdal-bin python-gdal
```

Note that this may affect previously installed versions of the _rgdal_, _sf_, and _maptools R_ packages. If these _R_ packages---or indeed any other packages---fail to load after running the previous system commands, you will need to reinstall the _R_ packages to use them.

### Example usage

The _tiler R_ package is designed to be used in conjunction with the _leaflet R_ package. Below are some examples using the _tiler R_ package.

## Leaflet

Here is an example of using the tiler package to visualize a small raster data set. Although the data set used in this example is rather small, the `tiles` function can be used to create tile data for large raster data sets. Note that you will need to use `options = tileOptions(tms = TRUE)` when adding tiles to a map. If you forget to include this option, then the map will not show the spatial data.

```{r, eval = FALSE}
# load packages
library(tiler)
library(leaflet)

# load example data set
data(netherlands, package = "mapmisc")

# create leaflet map and visualize data using tiles
map <- leaflet() %>%
       setView(lng = 5.75560, lat = 50.94723, zoom = 12) %>%
       addProviderTiles("Esri.WorldImagery", group = "Basemap") %>%
       addTiles(tiles(nldElev, output_dir = tempdir(), zoom = "10-16"),
                group = "Elevation",
                options = tileOptions(minZoom = 5, maxZoom = 16,
                                      tms = TRUE)) %>%
       addLayersControl(baseGroups = "Basemap", overlayGroups = "Elevation",
                        options = layersControlOptions(collapsed = FALSE))

# render map
print(map)
```

## Shiny

Now let's build a _shiny_ application that displays the spatial data set. Note that we have set `shiny = TRUE` to tell the `tiles` function that the tiles function that we are running it inside a _shiny_ application. Also, we have put the `tiles` function in a `start` function so the tiles are made while initializing the application (equivalent to putting code in a "global.R" _R shiny_ script).

```{r, eval = FALSE}
# load packages
library(shiny)
library(leaflet)
library(tiler)

# define function to build tiles on start
start <- function() {
  data(netherlands, package = "mapmisc")
  path <<- tiles(nldElev, output_dir = tempdir(), zoom = "10-16", shiny = TRUE)
}

# define ui
ui <- fluidPage(leafletOutput("map"))

# define server function
server <- function(input, output, session) {
  output$map <- renderLeaflet({
    leaflet() %>%
    setView(lng = 5.75560, lat = 50.94723, zoom = 12) %>%
    addProviderTiles("Esri.WorldImagery", group = "Basemap") %>%
    addTiles(path, group = "Elevation",
             options = tileOptions(minZoom = 5, maxZoom = 16, tms = TRUE)) %>%
    addLayersControl(baseGroups = "Basemap", overlayGroups = "Elevation",
                     options = layersControlOptions(collapsed = FALSE))
  })
}

# launch shiny app
shinyApp(ui, server, onStart = start)
```
